---
title: "Mobile Phone Trends"
author: "Carl Howe, RStudio"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googlesheets)
library(googledrive)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Import Data 

We can read in the Shortest Mobile Phone Survey data that we just collected using the `googlesheets` package. This will take a moment or two to run.

```{r import, echo=FALSE, message=FALSE}

gs_deauth()
access_token <- gs_url("https://docs.google.com/spreadsheets/d/1aPEgsl_q332zYVIN2rsq5O6lZCLNsXW4mQlORhW3Hto/")
access_token <- gs_key("1aPEgsl_q332zYVIN2rsq5O6lZCLNsXW4mQlORhW3Hto")
responses <- gs_read(access_token)
```

We now have the data in a data frame called `responses` and it has `r nrow(responses)` survey responses in it.

## Tidying the Data

Let's look at the first 5 rows of the data to get a sense of how it is formatted.

```{r tidy, echo=FALSE, message=FALSE}
head(responses, 5)
```

This data looks pretty tidy (we'll define what that means later), so we don't have to do anything in this step.

## Data Transformation

We'll summarize the results to prepare them for visualization. We'll just count the proportion of Apple responses in each column to get a sense for how Apple ownership has changed in the latest phone upgrade cycle and see how it might change in the future.

```{r transform, echo=FALSE, message=FALSE}
apple_summary <- responses %>% 
  summarize(Past = sum(PreviousPhone == "Apple")/n(),
            Present = sum(CurrentPhone == "Apple")/n(),
            Future = sum(NextPhone == "Apple")/n()
  ) %>% 
  pivot_longer(Past:Future, values_to = "Percent", names_to = "When") %>% 
  mutate(When = factor(When, levels=c("Past", "Present", "Future")))
print(apple_summary)

```

## Visualizing The Survey Data

We'll visualize the data using a chart that with points to show the proportions of Apple users in each ownership cycle.

```{r visualize, echo=FALSE, message=FALSE}
theme_set(theme_minimal())
ggplot(apple_summary, 
       aes(x = When,
           y = Percent)) +
     geom_point() +
     scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
     labs(y = "Percent Apple Owners", x = "")
```

## Modeling The Results

We can understand the overall trend by fitting a linear model to our 3 data points. This is quantitatively meaningless because our times are imprecise, but the linear model gives us a feel for what's going on with Apple device ownership and where it is headed in the future.

```{r model, echo=FALSE, message=FALSE}
ggplot(apple_summary, 
       aes(x = When,
           y = Percent)) +
     geom_point() +
     scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
     labs(y = "Percent Apple Owners", x = "") +
     geom_smooth(aes(x = as.numeric(When)), method = lm, se = FALSE)
```

