---
title: "Data Basics"
output: html_document
---

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License. -->

# R Packages

```{r setup}
library(tidyverse)
library(gapminder)
library(readxl)
```

## Your Turn 1

What R packages are being loaded in the first chunk of 02-Data-Basics.Rmd?

# Tabular Data

## Your Turn 2

Take a look at the `mpg` dataset in two ways:

1. Run the code in this chunk:
```{r}
mpg
```

2. Type mpg on the "Console" and hit Enter

What do you notice about the difference in the way the data is displayed?

## Your Turn 3

Try running each line one by one, with Ctrl + Enter (or Cmd + Enter on Mac):

```{r}
dim(x = mpg)     
names(x = mpg)   
glimpse(x = mpg) 
View(x = mpg)  
```

## Your Turn 4

How many rows of data are in `gapminder`?

```{r}

```

What are the names of the variables in `gapminder`?

```{r}

```

## Your Turn 5

Run the chunk to get help on `mpg`. What is this data?
```{r}
?mpg
```

# Vector Data

## Your Turn 6

Take another look at mpg.  What kind of data is in each column?

```{r}
mpg
```

# Importing Data

```{r}
nimbus <- read_csv("data/nimbus.csv")
```

## Your Turn 7

What code do you need to read in deaths.xls?

Use the Import Data tool to help generate the code.

```{r}


```

*Challenge:* Can you see a problem in the imported data? Can you Import again and fix it?

```{r}

```


