---
title: "Data Basics"
output: html_notebook
---

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License. -->

```{r setup}
library(tidyverse)
library(haven)
library(readxl)
```

# Tabular Data

You might not see much difference between these two in this notebook, but one is a tibble and one is a data frame:
```{r}
mpg
```

```{r}
quakes
```

## Your turn 1

Take a look at these two datasets, by typing their names into the console:
* quakes
* mpg

## Your turn 2

Try running each line one by one, with Cmd/Ctrl + r:

```{r}
dim(x = mpg)     
names(x = mpg)   
glimpse(x = mpg) 
View(x = mpg)  
```

## Your turn 3

How many rows of data are in quakes?

```{r}

```

What are the names of the variables in quakes?

```{r}

```

## Your turn 4

Run the chunk to get help on `mpg`. What is this data?
```{r}
?mpg
```

# Vector Data

## Your turn 5

Take another look at mpg.  What kind of data is in each column?

```{r}
mpg
```

# Importing Data

## Your Turn 6

Take a look in the data/ directory in the project folder.

Try reading in each file with the appropriate function.

```{r}

```


## Your Turn 7

Can you see what is wrong with the Excel file when it is imported?

Scan the Arguments section of ?read_excel, can you find an argument that might help? Try it!

```{r}
?read_excel
```

