---
title: "Transform Data"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(gapminder)

# Toy dataset to use
pollution <- tribble(
       ~city,   ~size, ~amount, 
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",      121,
   "Beijing", "small",      56
)
```

## gapminder

```{r}
gapminder
```

## Your Turn 1

* `filter()` selects rows
* logical tests

See if you can use the logical operators to manipulate our code below to show:

The data for Canada
```{r}
filter(gapminder, country == "New Zealand")
```

All data for countries in Oceania
```{r}
filter(gapminder, country == "New Zealand")
```

Rows where the life expectancy is greater than 82
```{r}
filter(gapminder, country == "New Zealand")
```


## Your Turn 2

Use Boolean operators to alter the code below to return only the rows that contain:

* Canada before 1970
```{r}
filter(gapminder, country == "Canada", year < 1970)
```

*  Countries where life expectancy in 2007 is below 50

```{r}
filter(gapminder, year == 2007, lifeExp < 50)
```

* Countries where life expectancy in 2007 is below 50, and are not in Africa.

```{r}
filter(gapminder, year == 2007, lifeExp < 50, !(continent == "Africa"))
```

## Your Turn 3

Alter the code to:

* Add an africa column, which contains TRUE is the country is on the Africa continent.
* Add a rank_pop column to rank each row in gapminder from largest pop to smallest pop.


```{r}
mutate(gapminder, 
  africa = continent == "Africa",
  rank_pop = min_rank(desc(pop)))
```

## Your Turn 4

Use summarise() to compute three statistics about the data:

* The first (minimum) year in the dataset
* The last (maximum) year in the dataset
* The number of countries represented in the data (Hint: use cheatsheet)

```{r}
gapminder %>% 
  summarise(first = min(year), 
            last = max(year),
            n_countries = n_distinct(country))
```

## Your Turn 5

Extract the rows where continent == "Africa" and year == 2007. 

Then use summarise() and summary functions to find:

1. The number of unique countries
2. The median life expectancy

```{r}
gapminder %>% 
  filter(continent == "Africa", year == 2007) %>% 
  summarise(n_countries = n_distinct(country),
    med_life_exp = median(lifeExp))
```


```{r}
gapminder %>% 
  group_by(continent)
```
## Your Turn 6

Find the median life expectancy by continent

```{r}
gapminder %>%
  group_by(continent) %>% 
  summarise(med_life_exp = median(lifeExp)) 
```

## Your Turn 7

Brainstorm with your neighbor the sequence of operations to find:  the country with biggest jump in life expectancy  (between any two consecutive records) for each continent.

1. Find jumps between time points for all countries:

    i) Group by country
    ii) Add jump variable
2. Rank jump within continent

    i) Group by continent
    ii) Add rank of jump variable
3. Filter out countries with rank 1

## Your Turn 8

Find the country with biggest jump in life expectancy (between any two consecutive records) for each continent.

```{r}
# One of many solutions
gapminder %>%
  group_by(country) %>%
  mutate(jump = lifeExp - lag(lifeExp)) %>%
  group_by(continent) %>%
  mutate(rank = min_rank(desc(jump))) %>%
  filter(rank == 1)

```

***

# Take aways

* Extract cases with `filter()`  
* Make new variables, with `mutate()`  
* Make tables of summaries with `summarise()`  
* Do groupwise operations with `group_by()`
* Connect operations with `%>%`  
