---
title: "Modelling & Broom"
output: html_notebook
---

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License. -->

```{r setup}
library(gapminder)
library(tidyverse)
library(broom)
```

## Your turn 1 

Run the code chunk.  

Take another look at life expectancy over time for Canada.  What do you notice? 

```{r}
canada <- gapminder %>%
  filter(country == "Canada")

ggplot(canada) +
  geom_line(mapping = aes(x = year, y = lifeExp))
```

## Your turn 2

Run the chunk to fit the model.
What kind of output do you get?
Explore canada_lm by clicking on it in the environment, what's inside the object?

```{r}
canada_lm <- lm(lifeExp ~ year, data = canada)
canada_lm
```

## Your turn 3

Look at the output of canada_lm %>% glance()
Use dplyr tools to extract the R-squared for the model.

```{r}
canada_lm %>% glance() 
```


## Your turn 4

How has life expectancy changed in other countries?
Make a line plot of lifeExp vs. year grouped by country.  
Set alpha to 0.2, to see the results better.

```{r}
ggplot(gapminder)


```

