---
title: "Iteration and List Columns"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(gapminder)
library(broom)

nz <- gapminder %>%
  filter(country == "New Zealand")
canada <- gapminder %>%
  filter(country == "Canada")
us <- gapminder %>%
  filter(country == "United States")

# A complicated object to play with
three_models <- list(
  lm(lifeExp ~ year, data = nz),
  lm(lifeExp ~ year, data = canada),
  lm(lifeExp ~ year, data = us)
)
```

# Iteration

## Your turn 1

Run the setup chunk.
Look for  three_models in your environment.  Click on it to open a Viewer.
What kind of object is this?  Does its contents look familiar?

```{r}

```

## Your Turn 2

Do it for one element: add some code to tidy nz_model.

```{r}
nz_model <- three_models[[1]]

```

## Your turn 3

Run the code, what kind of object is `tidy_models`?

```{r}
tidy_models <- map(three_models, ~ tidy(.x))
tidy_models
```

## Your turn 4

Edit the code to instead get the r.squared for each model.

```{r}
map_dbl(three_models, ~ tidy(.x) %>% 
                       filter(term == "year") %>% 
                       pull(estimate)
)
```

# List columns


## Your turn 5

Run the chunk: it nests gapminder, then pulls out the data column into data_column.
Use the object explorer (i.e. the magnifying glass in the Environment pane)mto take a look at `data_column`.  What kind of object is it?

```{r}
gapminder_nested <- gapminder %>% 
  group_by(country) %>% 
  nest()

data_column <- gapminder_nested %>% pull(data) 
```

Following along in slides:
```{r}
# Add the models to a model column by mapping over data
gapminder_nested <- gapminder_nested %>% 
  mutate(model = map(data, ~ lm(lifeExp ~ year, data = .x)))

# Create the r-squared column, by mapping over model
gapminder_nested <- gapminder_nested %>% 
  mutate(r.squared = map_dbl(model, 
                        ~ glance(.x) %>% pull(r.squared)))
```

## Your turn 6

(Make sure you run all the chunks before this one, then)
Filter `gapminder_nested` to find the countries with r.squared less than 0.5.  

```{r}
gapminder_nested %>% filter(r.squared < 0.5)
```

Following along with slides:
```{r}
gapminder_nested %>% 
  filter(r.squared < 0.5) %>%
  unnest(data) %>%
  ggplot(aes(x = year, y = lifeExp)) + 
    geom_line(aes(color = country))

```

## Your Turn 7

Write code to:

* Add a column to `gapminder_nested` with the slope coefficient from the model
* Filter to find countries with a slope above 0.6 years/year.
* Plot these countries over time.

```{r}
gapminder_nested 
```
